---
title: "diffusion_analysis_shared"
author: "Hanna Willis"
date: "15/02/2024"
output: html_document
---
# ------------ LOAD PACKAGES ------------
```{r}
library(ggplot2)
library(emmeans)
```

# ------------ LOAD DATA ------------
```{r, include=FALSE}
# Behavioural results
ppt_details_change <- read.csv('all_behaviour_Feb24.csv')

# All diffusion
all_diff <- read.csv("/Users/hwillis_admin/Desktop/PostDocs/Projects/Rehab_Project/MRI/Diffusion/data/sharing/share/mean_blind_tractmeasures.csv")

# Diffusion change
change_all <- read.csv("change_all.csv")
change_all_distal <- read.csv("change_all_distal.csv")

# Full tract for hMT+ 
hMT_fulltract_data_trained_only <- read.csv("hMT_fulltract_data_trained_only.csv")

# Distal dLGN for all timepoints 
hMT_distal_data_alltime <- read.csv("hMT_distal_data_alltime.csv")

# Distal dLGN in sighted field after training
combined_df <- read.csv("combined_df.csv")
```

# ------------ MANIPULATE DATAFRAMES ------------
```{r}
# Set up factor levels 
hMT_fulltract_data_trained_only$Timepoint <- factor(hMT_fulltract_data_trained_only$Timepoint, levels = c("Pre-Baseline", "Pre-Training", "Post-Training"))
hMT_distal_data_alltime$Timepoint <- factor(hMT_distal_data_alltime$Timepoint, levels = c("Pre-Baseline", "Pre-Training", "Post-Training"))
combined_df$Timepoint <- factor(combined_df$Timepoint, levels = c("Pre-Baseline", "Pre-Training", "Post-Training"))

# Manipulate dataframes
ppt_details_change_MRI <- ppt_details_change[!grepl("R016|R021", ppt_details_change$Subj),]
hMT_distal_data_trained_only_sight <- combined_df[grepl("sight", combined_df$hemisphere),]
hMT_distal_data_trained_only <- combined_df[grepl("blind", combined_df$hemisphere),]

# Include only ppt who have both timepoints
hMT_distal_data_pretrain <- hMT_distal_data_alltime %>%
  group_by(subjectID) %>%
  filter(all(c("Pre-Baseline", "Pre-Training") %in% Timepoint))

## Remove Post-Training visit
hMT_distal_data_pretrain_only <- hMT_distal_data_pretrain[!grepl("Post-Training", hMT_distal_data_pretrain$Timepoint),]


# Shouldn't need these but check 

#hMT_distal_data_trained_only$hemisphere <- "blind"
#hMT_distal_data_trained_only_sight$hemisphere <- "sight"
#combined_df <- rbind(hMT_distal_data_trained_only_sight, hMT_distal_data_trained_only)
#hMT_distal_data_trained_only_sight <- read.csv("/Users/hwillis_admin/Desktop/PostDocs/Projects/Rehab_Project/MRI/Diffusion/data/sharing/hMT_distal_data_trained_only_sight.csv")
#hMT_distal_data_trained_only <- read.csv("/Users/hwillis_admin/Desktop/PostDocs/Projects/Rehab_Project/MRI/Diffusion/data/sharing/hMT_distal_data_trained_only.csv")

```

# ------------ MAKE FIGURES ------------

### Figure 4: Relationship between mean FA and behaviour
```{r}
# Motion discrimination 
# ---------------------
p3 <- ggplot(change_all, aes(x = change_motion_improve_ndr_pos, y = change_fa_hMT)) +
  geom_point(size = 3, alpha = 0.4,colour="orange1") +
  labs(x = "Improvement in \nNDR thresholds (%)", y= "Mean FA change") +
  ggtitle("dLGN-hMT+") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(-0.05,0.05), breaks = seq(from=-0.04, to=0.04, by=0.02)) +
  theme(plot.title = element_text(hjust = 0.5, face="bold",size=18)) 

p2 <- ggplot(change_all, aes(x = change_motion_improve_ndr_pos, y = change_fa_v1)) +
  geom_point(size = 3, alpha = 0.4,colour="#CC3399") +
  labs(x = "Improvement in \nNDR thresholds (%)", y= "Mean FA change") +
  ggtitle("dLGN-V1") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(-0.05,0.05), breaks = seq(from=-0.04, to=0.04, by=0.02)) +
  theme(plot.title = element_text(hjust = 0.5, face="bold",size=18)) 

p1 <- ggplot(change_all, aes(x = change_motion_improve_ndr_pos, y = change_fa_motor)) +
  geom_point(size = 3, alpha = 0.4,colour="dodgerblue") +
  labs(x = "Improvement in \nNDR thresholds (%)", y= "Mean FA change") +
  ggtitle("VPL-S1") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(-0.05,0.05), breaks = seq(from=-0.04, to=0.04, by=0.02)) +
  theme(plot.title = element_text(hjust = 0.5, face="bold",size=18)) 

# HVF area improved
# ---------------------
p3 <- ggplot(change_all, aes(x = area_improve_hvf, y = change_fa_hMT)) +
  geom_point(size = 3, alpha = 0.4,colour="orange1") +
  labs(y= "Mean FA change", x = expression("Area Improved (" * deg^2 * ")")) +
  ggtitle("dLGN-hMT+") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(-0.05,0.05), breaks = seq(from=-0.05, to=0.05, by=0.02)) +
  theme(plot.title = element_text(hjust = 0.5, face="bold",size=18)) 

## --------- V1 ---------
p2 <- ggplot(change_all, aes(x = area_improve_hvf, y = change_fa_v1)) +
  geom_point(size = 3, alpha = 0.4,colour="#CC3399") +
  labs(y= "Mean FA change", x = expression("Area Improved (" * deg^2 * ")")) +
  ggtitle("dLGN-V1") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(-0.05,0.05), breaks = seq(from=-0.05, to=0.05, by=0.02)) +
  theme(plot.title = element_text(hjust = 0.5, face="bold",size=18)) 

## --------- VPL ---------
p1 <- ggplot(change_all, aes(x = area_improve_hvf, y = change_fa_motor)) +
  geom_point(size = 3, alpha = 0.4,colour="dodgerblue") +
  labs(y= "Mean FA change", x = expression("Area Improved (" * deg^2 * ")")) +
  ggtitle("VPL-S1") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(-0.05,0.05), breaks = seq(from=-0.05, to=0.05, by=0.02)) +
  theme(plot.title = element_text(hjust = 0.5, face="bold",size=18)) 
```

### Figure 5: The distal dLGN-hMT+ tract (pre-post training) 
```{r}
# 1) FA along full tract
# ----------------------

# Calculate SEM
hMT_full_tract_data_trained_fig <- hMT_fulltract_data_trained_only %>%
  group_by(nodeID, Timepoint) %>%
  summarise(Mean_FA = mean(fa), SEM = sd(fa) / sqrt(length(fa)))

# Create a ggplot line plot with a smooth line and SEM error ribbons
p1 <- ggplot(hMT_full_tract_data_trained_fig, aes(x = nodeID, y = Mean_FA, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_FA - SEM, ymax = Mean_FA + SEM, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average FA") +
  ggtitle("Full dLGN-hMT+\n") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.23,0.50), breaks = seq(from=0.25, to=0.50, by=0.05)) +
  scale_x_continuous(limits = c(15,85), breaks = seq(from=15, to=90, by=10)) +
  theme(legend.position = c(0.7, 0.9)) +
  theme(plot.title = element_text(hjust = 0.5)) 
p1

# 2) diffusion metrics in distal tract 
# ----------------------

# Calculate the standard error of the mean (SEM) for each node and timepoint
hMT_distal_data_trained_fig <- hMT_distal_data_trained_only %>%
  group_by(nodeID, Timepoint) %>%
  summarise(Mean_FA = mean(fa), SEM_FA = sd(fa) / sqrt(length(fa)),
            Mean_MD = mean(md), SEM_MD = sd(md) / sqrt(length(md)),
            Mean_NDI = mean(ndi), SEM_NDI = sd(ndi) / sqrt(length(ndi)),
            Mean_ODI = mean(odi), SEM_ODI = sd(odi) / sqrt(length(odi)),
            Mean_ISOVF = mean(isovf), SEM_ISOVF = sd(isovf) / sqrt(length(isovf)))

p1 <- ggplot(hMT_distal_data_trained_fig, aes(x = nodeID, y = Mean_FA, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_FA - SEM_FA, ymax = Mean_FA + SEM_FA, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average FA") +
  ggtitle("FA") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.23,0.40), breaks = seq(from=0.25, to=0.40, by=0.05)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "top") +
  theme(plot.title = element_text(hjust = 0.5)) 
p1

p2 <- ggplot(hMT_distal_data_trained_fig, aes(x = nodeID, y = Mean_MD, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_MD - SEM_MD, ymax = Mean_MD + SEM_MD, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average MD") +
  ggtitle("MD") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.80,1.10), breaks = seq(from=0.80, to=1.10, by=0.1)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p2

p3 <- ggplot(hMT_distal_data_trained_fig, aes(x = nodeID, y = Mean_NDI, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_NDI - SEM_NDI, ymax = Mean_NDI + SEM_NDI, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average NDI") +
  ggtitle("NDI") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.45,0.60), breaks = seq(from=0.45, to=0.60, by=0.05)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p3

p4 <- ggplot(hMT_distal_data_trained_fig, aes(x = nodeID, y = Mean_ODI, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_ODI - SEM_ODI, ymax = Mean_ODI + SEM_ODI, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average ODI") +
  ggtitle("ODI") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.14,0.32), breaks = seq(from=0.14, to=0.32, by=0.06)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "top") +
  theme(plot.title = element_text(hjust = 0.5)) 
p4

p5 <- ggplot(hMT_distal_data_trained_fig, aes(x = nodeID, y = Mean_ISOVF, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_ISOVF - SEM_ISOVF, ymax = Mean_ISOVF + SEM_ISOVF, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average ISOVF") +
  ggtitle("ISOVF") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.02,0.08), breaks = seq(from=0.02, to=0.08, by=0.02)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p5

```

### Figure 6: The distal dLGN-hMT+ tract (pre-baseline-pre-training)
```{r}

# Calculate the standard error of the mean (SEM) for each node and timepoint
hMT_distal_data_pretrain_fig <- hMT_distal_data_pretrain_only %>%
  group_by(nodeID, Timepoint) %>%
  summarise(Mean_FA = mean(fa), SEM_FA = sd(fa) / sqrt(length(fa)),
            Mean_MD = mean(md), SEM_MD = sd(md) / sqrt(length(md)),
            Mean_NDI = mean(ndi), SEM_NDI = sd(ndi) / sqrt(length(ndi)),
            Mean_ODI = mean(odi), SEM_ODI = sd(odi) / sqrt(length(odi)),
            Mean_ISOVF = mean(isovf), SEM_ISOVF = sd(isovf) / sqrt(length(isovf)))

p1 <- ggplot(hMT_distal_data_pretrain_fig, aes(x = nodeID, y = Mean_FA, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_FA - SEM_FA, ymax = Mean_FA + SEM_FA, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("grey","#FFCC33"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average FA") +
  ggtitle("FA") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.23,0.40), breaks = seq(from=0.25, to=0.40, by=0.05)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p1

p2 <- ggplot(hMT_distal_data_pretrain_fig, aes(x = nodeID, y = Mean_MD, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_MD - SEM_MD, ymax = Mean_MD + SEM_MD, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("grey","#FFCC33"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average MD") +
  ggtitle("MD") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.80,1.10), breaks = seq(from=0.80, to=1.10, by=0.1)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p2

p3 <- ggplot(hMT_distal_data_pretrain_fig, aes(x = nodeID, y = Mean_NDI, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_NDI - SEM_NDI, ymax = Mean_NDI + SEM_NDI, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("grey","#FFCC33"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average NDI") +
  ggtitle("NDI") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.45,0.60), breaks = seq(from=0.45, to=0.60, by=0.05)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p3

p4 <- ggplot(hMT_distal_data_pretrain_fig, aes(x = nodeID, y = Mean_ODI, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_ODI - SEM_ODI, ymax = Mean_ODI + SEM_ODI, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("grey","#FFCC33"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average ODI") +
  ggtitle("ODI") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.14,0.32), breaks = seq(from=0.14, to=0.32, by=0.06)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p4

p5 <- ggplot(hMT_distal_data_pretrain_fig, aes(x = nodeID, y = Mean_ISOVF, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_ISOVF - SEM_ISOVF, ymax = Mean_ISOVF + SEM_ISOVF, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("grey","#FFCC33"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average ISOVF") +
  ggtitle("ISOVF") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0,0.12), breaks = seq(from=0, to=0.12, by=0.04)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p5

```

### Figure 7: Sighted dLGN-hMT+ (supplementary materials)
```{r}
# Calculate the standard error of the mean (SEM) for each node and timepoint
hMT_distal_data_trained_fig_sight <- hMT_distal_data_trained_only_sight %>%
  group_by(nodeID, Timepoint) %>%
  summarise(Mean_FA = mean(fa), SEM_FA = sd(fa) / sqrt(length(fa)),
            Mean_MD = mean(md), SEM_MD = sd(md) / sqrt(length(md)),
            Mean_NDI = mean(ndi), SEM_NDI = sd(ndi) / sqrt(length(ndi)),
            Mean_ODI = mean(odi), SEM_ODI = sd(odi) / sqrt(length(odi)),
            Mean_ISOVF = mean(isovf), SEM_ISOVF = sd(isovf) / sqrt(length(isovf)))

p1 <- ggplot(hMT_distal_data_trained_fig_sight, aes(x = nodeID, y = Mean_FA, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_FA - SEM_FA, ymax = Mean_FA + SEM_FA, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average FA") +
  ggtitle("FA") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.23,0.40), breaks = seq(from=0.25, to=0.40, by=0.05)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "top") +
  theme(plot.title = element_text(hjust = 0.5)) 
p1

p2 <- ggplot(hMT_distal_data_trained_fig_sight, aes(x = nodeID, y = Mean_MD, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_MD - SEM_MD, ymax = Mean_MD + SEM_MD, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average MD") +
  ggtitle("MD") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.80,1.10), breaks = seq(from=0.80, to=1.10, by=0.1)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p2

p3 <- ggplot(hMT_distal_data_trained_fig_sight, aes(x = nodeID, y = Mean_NDI, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_NDI - SEM_NDI, ymax = Mean_NDI + SEM_NDI, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average NDI") +
  ggtitle("NDI") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.45,0.65), breaks = seq(from=0.45, to=0.65, by=0.05)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p3

p4 <- ggplot(hMT_distal_data_trained_fig_sight, aes(x = nodeID, y = Mean_ODI, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_ODI - SEM_ODI, ymax = Mean_ODI + SEM_ODI, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average ODI") +
  ggtitle("ODI") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.14,0.32), breaks = seq(from=0.14, to=0.32, by=0.06)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p4

p5 <- ggplot(hMT_distal_data_trained_fig_sight, aes(x = nodeID, y = Mean_ISOVF, color = Timepoint, group = Timepoint)) +
  geom_line(size = 2) +
  geom_ribbon(aes(ymin = Mean_ISOVF - SEM_ISOVF, ymax = Mean_ISOVF + SEM_ISOVF, fill = Timepoint), alpha = 0.3, color = NA) +
  scale_fill_manual(values = c("#FFCC33","orange1"), aesthetics = c("color", "fill")) +
  labs(x = "Node", y= "Average ISOVF") +
  ggtitle("ISOVF") +
  theme_classic(base_size = 18) +
  scale_y_continuous(limits = c(0.02,0.10), breaks = seq(from=0.02, to=0.10, by=0.02)) +
  scale_x_continuous(limits = c(60,85), breaks = seq(from=60, to=85, by=10)) +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5)) 
p5

```

# ------------ RUN STATS ------------

# Behaviour change
```{r}
# Motion discrimination
# ---------------------
## Perc correct
mean(ppt_details_change_MRI$pre_motion_improve_perc)
sd(ppt_details_change_MRI$pre_motion_improve_perc)
mean(ppt_details_change_MRI$post_motion_improve_perc)
sd(ppt_details_change_MRI$post_motion_improve_perc)
t.test(ppt_details_change_MRI$pre_motion_improve_perc, ppt_details_change_MRI$post_motion_improve_perc, paired = TRUE)

## NDR
mean(ppt_details_change_MRI$pre_motion_improve_ndr)
sd(ppt_details_change_MRI$pre_motion_improve_ndr)
mean(ppt_details_change_MRI$post_motion_improve_ndr)
sd(ppt_details_change_MRI$post_motion_improve_ndr)
t.test(ppt_details_change_MRI$pre_motion_improve_ndr, ppt_details_change_MRI$post_motion_improve_ndr, paired = TRUE)

# HVF
# ---------------------

## Area improved (R002 and R014 removed due to errors outside norm)
ppt_details_changexR002xR014 <- ppt_details_change_MRI[!grepl("R002|R014", ppt_details_change_MRI$Subj),]
mean(ppt_details_changexR002xR014$area_improve_hvf)
sd(ppt_details_changexR002xR014$area_improve_hvf)
range(ppt_details_changexR002xR014$area_improve_hvf)
t.test(ppt_details_changexR002xR014$area_improve_hvf, mu = 0)

# Relationship between behaviour
# ---------------------
model <- lm(change_motion_improve_ndr~area_improve_hvf, data=ppt_details_changexR002xR014)
summary(model)

```

# Mean diffusion changes
```{r}
## VPL
  ## BASIC STATS
  basic_stats <- group_by(vpl, sessionID) %>%
      summarise(
        count = n(),
        FA_median = median(average_FA, na.rm = TRUE),
        FA_IQR = IQR(average_FA, na.rm = TRUE),
      )

## V1
  ## BASIC STATS
  basic_stats <- group_by(v1, sessionID) %>%
      summarise(
        count = n(),
        median = median(average_FA, na.rm = TRUE),
        IQR = IQR(average_FA, na.rm = TRUE)
      )

## hMT+
  ## BASIC STATS
  basic_stats <- group_by(hMT, sessionID) %>%
      summarise(
        count = n(),
        median = median(average_FA, na.rm = TRUE),
        IQR = IQR(average_FA, na.rm = TRUE)
      )
  
```


# Mean diffusion metrics and behaviour
```{r} 
# Create an empty data frame to store regression results
regression_results <- data.frame(
  Independent_Variable = character(),
  Dependent_Variable = character(),
  Adjusted_R2 = numeric(),
  F_Stat = numeric(),
  Degrees_of_Freedom = integer(),
  P_Value = numeric()
)

# Loop through each independent variable you want to regress
independent_variables <- c("change_fa_v1", "change_fa_motor","change_fa_hMT","change_md_hMT", "change_rd_hMT", "change_ad_hMT","change_ndi_hMT","change_odi_hMT","change_isovf_hMT","baseline_fa_hMT","baseline_fa_v1","baseline_fa_motor")


# Loop through each dependent variable
dependent_variables <- c("change_motion_improve_ndr_pos", "area_improve_hvf")
  
for (dependent_var in dependent_variables) {
  for (independent_var in independent_variables) {
  # Perform linear regression
  lm_model <- lm(paste(dependent_var, "~", independent_var, sep = ""), data = change_all)

  # Extract regression results and store them in a data frame
  results <- tidy(lm_model)
  
  # Calculate adjusted R-squared
  adjusted_r2 <- summary(lm_model)$adj.r.squared
  
  # Calculate degrees of freedom
  df <- summary(lm_model)$fstatistic[3]
  f_stat <- summary(lm_model)$fstatistic[1]
    
  # Calculate p-value
  p_value <- summary(lm_model)$coefficients[2,4]
  
  # Add the dependent variable name and other results to the table
  regression_results <- rbind(
    regression_results,
    data.frame(
      Independent_Variable = independent_var,
      Dependent_Variable = dependent_var,
      Adjusted_R2 = adjusted_r2,
      F_Stat = f_stat,
      Degrees_of_Freedom = df,
      P_Value = p_value
    )
  )
}}

# Print the table with regression results
print(regression_results)

# To run without R011 
## Without R011 from HVF - run again with this dataframe
change_all_xR011 <- change_all[!grepl("R011", change_all$Subj),]

```

# Changes along the tract with training
```{r}
# FA
# ------------
# Run LME
hMT_blind_fa_distal_model <- lmer(fa ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_trained_only)
summary(hMT_blind_fa_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_fa_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

# MD 
# ------------
# Run LME
hMT_blind_md_distal_model <- lmer(md ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_trained_only)
summary(hMT_blind_md_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_md_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

# NDI
# ------------
# Run LME
hMT_blind_ndi_distal_model <- lmer(ndi ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_trained_only)
summary(hMT_blind_ndi_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_ndi_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

# ODI
# ------------
# Run LME
hMT_blind_odi_distal_model <- lmer(odi ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_trained_only)
summary(hMT_blind_odi_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_odi_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

# ISOVF 
# ------------
# Run LME
hMT_blind_isovf_distal_model <- lmer(isovf ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_trained_only)
summary(hMT_blind_isovf_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_isovf_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

```

# Changes along the tract without training
```{r}
## FA 
# ------------
# Run LME
hMT_blind_fa_distal_model <- lmer(fa ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_pretrain_only)
summary(hMT_blind_fa_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_fa_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

## MD
# ------------
hMT_blind_md_distal_model <- lmer(md ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_pretrain_only)
summary(hMT_blind_md_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_md_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

## NDI
# ------------
hMT_blind_ndi_distal_model <- lmer(ndi ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_pretrain_only)
summary(hMT_blind_ndi_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_ndi_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

## ODI
# ------------
hMT_blind_odi_distal_model <- lmer(odi ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_pretrain_only)
summary(hMT_blind_odi_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_odi_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

## ISOVF
# ------------
hMT_blind_isovf_distal_model <- lmer(isovf ~ 1 + Timepoint + nodeID +
                      (1|subjectID), 
                    data = hMT_distal_data_pretrain_only)
summary(hMT_blind_isovf_distal_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_isovf_distal_model, specs = ~Timepoint + nodeID)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

```

# Distal dLGN metrics and changes in behaviour
```{r}
# Create an empty data frame to store regression results
regression_results <- data.frame(
  Independent_Variable = character(),
  Dependent_Variable = character(),
  Adjusted_R2 = numeric(),
  F_Stat = numeric(),
  Degrees_of_Freedom = integer(),
  P_Value = numeric()
)

# Loop through each independent variable you want to regress
independent_variables <- c("change_fa_hMT","change_md_hMT", "change_rd_hMT", "change_ad_hMT","change_ndi_hMT","change_odi_hMT","change_isovf_hMT")


# Loop through each dependent variable
dependent_variables <- c("change_motion_improve_ndr_pos", "area_improve_hvf")
  
for (dependent_var in dependent_variables) {
  for (independent_var in independent_variables) {
  # Perform linear regression
  lm_model <- lm(paste(dependent_var, "~", independent_var, sep = ""), data = change_all_distal)

  # Extract regression results and store them in a data frame
  results <- tidy(lm_model)
  
  # Calculate adjusted R-squared
  adjusted_r2 <- summary(lm_model)$adj.r.squared
  
  # Calculate degrees of freedom
  df <- summary(lm_model)$fstatistic[3]
  f_stat <- summary(lm_model)$fstatistic[1]
    
  # Calculate p-value
  p_value <- summary(lm_model)$coefficients[2,4]
  
  # Add the dependent variable name and other results to the table
  regression_results <- rbind(
    regression_results,
    data.frame(
      Independent_Variable = independent_var,
      Dependent_Variable = dependent_var,
      Adjusted_R2 = adjusted_r2,
      F_Stat = f_stat,
      Degrees_of_Freedom = df,
      P_Value = p_value
    )
  )
}}

# Print the table with regression results
print(regression_results)

```

# Sighted dLGN-hMT+ (supplementary materials)
```{r}
## FA 
# ------------
# Run LME
hMT_blind_fa_distal_model_combined <- lmer(fa ~ 1 + hemisphere * (nodeID + Timepoint) +
                      (1|subjectID), 
                      data = combined_df)
summary(hMT_blind_fa_distal_model_combined)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_fa_distal_model_combined, specs = ~Timepoint *hemisphere)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

## MD
# --------------
# Run LME
hMT_blind_md_distal_model_combined <- lmer(md ~ 1 + hemisphere * (nodeID + Timepoint) +
                      (1|subjectID), 
                      data = combined_df)
summary(hMT_blind_md_distal_model_combined)
report(hMT_blind_md_distal_model_combined)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_md_distal_model_combined, specs = ~Timepoint *hemisphere)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

## NDI
# --------------
# Run LME
hMT_blind_ndi_distal_model_combined <- lmer(ndi ~ 1 + hemisphere * (nodeID + Timepoint) +
                      (1|subjectID), 
                      data = combined_df)
summary(hMT_blind_ndi_distal_model_combined)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_ndi_distal_model_combined, specs = ~Timepoint *hemisphere)


# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

## ODI
# --------------
# Run LME
hMT_blind_odi_distal_model_combined <- lmer(odi ~ 1 + hemisphere * (nodeID + Timepoint) +
                      (1|subjectID), 
                      data = combined_df)
summary(hMT_blind_odi_distal_model_combined)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_odi_distal_model_combined, specs = ~Timepoint *hemisphere)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

## ISOVF
# --------------
# Run LME
hMT_blind_isovf_distal_model_combined <- lmer(isovf ~ 1 + hemisphere * (nodeID + Timepoint) +
                      (1|subjectID), 
                      data = combined_df)
summary(hMT_blind_isovf_distal_model_combined)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(hMT_blind_isovf_distal_model_combined, specs = ~Timepoint *hemisphere)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

# Adjust pvalues
p_values_diff <- c(.0001,.02,.24,.0004,.007,.0075,.0001,.5798,.9964,.043)
p.adjust(p_values_diff,method="holm") 

```

# Data quality checks (supplementary materials)
```{r}
# Mean FA in white matter
# -----------------------
mean_fa_wm <- mean_fa[grepl("white_matter", mean_fa$structureID),]

## Summary data
mean_fa_wm %>%
  group_by(timepoint) %>%
  get_summary_stats(fa_mean, type = "mean_sd")

## LME
mean_fa_model <- lmer(fa_mean ~ 1 + timepoint +
                      (1|subjectID), 
                    data = mean_fa_wm)
summary(mean_fa_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(mean_fa_model, specs = ~timepoint)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)


## SNR
# -----------------------

## Summary data
all_snr %>%
  group_by(timepoint) %>%
  get_summary_stats(snr, type = "mean_sd")

mean_fa_model <- lmer(snr ~ 1 + timepoint +
                      (1|subj), 
                    data = all_snr)
summary(mean_fa_model)

# Calculate EMMs and pairwise comparisons
emm <- emmeans(mean_fa_model, specs = ~timepoint)

# Perform pairwise comparisons, adjust for multiple comparisons (e.g., Tukey)
comparisons <- pairs(emm, adjust = "tukey")

# Display the results
print(comparisons)

# Tracts of interest
# -----------------------
# VPL
t_test(vpl_all_blind_tractmeasures, average_FA ~ sessionID, paired=TRUE)

# V1
t_test(v1_all_blind_tractmeasures, average_FA ~ sessionID, paired=TRUE)

# hMT
t_test(hMT_all_blind_tractmeasures, average_FA ~ sessionID, paired=TRUE)

```
